\documentclass[11pt,a4paper]{article}

\usepackage[margin=1in]{geometry}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{xcolor}
\usepackage{colortbl}
\usepackage{tikz}
\usepackage{enumitem}
\usepackage{amsmath}
\usepackage{hyperref}
\usepackage{float}
\usepackage{caption}
\usepackage{tabularx}

\usetikzlibrary{arrows.meta, positioning, shapes.geometric, calc}

\definecolor{yes}{HTML}{2E7D32}
\definecolor{no}{HTML}{C62828}
\definecolor{partial}{HTML}{F57F17}
\definecolor{degraded}{HTML}{E65100}
\definecolor{lightgray}{HTML}{F5F5F5}
\definecolor{sectionblue}{HTML}{1565C0}

\newcommand{\yes}{\textcolor{yes}{\textbf{Yes}}}
\newcommand{\no}{\textcolor{no}{\textbf{No}}}
\newcommand{\prt}{\textcolor{partial}{\textbf{Partial}}}
\newcommand{\dgr}{\textcolor{degraded}{\textbf{Degraded}}}
\newcommand{\na}{\textcolor{gray}{N/A}}

\newcolumntype{C}[1]{>{\centering\arraybackslash}m{#1}}
\newcolumntype{L}[1]{>{\raggedright\arraybackslash}m{#1}}

\title{\textbf{Bitchat Transport: Connectivity Analysis}\\[0.5em]
\large Address Types, Peer Topologies, and Mobility Scenarios}
\author{Bitchat Transport Layer Documentation}
\date{\today}

\begin{document}

\maketitle
\tableofcontents
\newpage

% =============================================================================
\section{Introduction}
% =============================================================================

This document provides a comprehensive analysis of connectivity between Bitchat peers across all combinations of network address types, peer topologies, and mobility scenarios. Bitchat uses two transport layers:

\begin{itemize}
  \item \textbf{BLE (Bluetooth Low Energy)} --- short-range, no Internet required, used for local peer discovery and communication
  \item \textbf{Iroh} --- QUIC-over-UDP P2P networking, uses relay servers for NAT traversal, supports direct connections when possible
\end{itemize}

\subsection{Key Constraints}

\begin{enumerate}
  \item \textbf{No store-and-forward}: If a peer is unreachable, the message fails. No caching, no queuing, no relaying through intermediaries.
  \item \textbf{No message relaying}: A third peer cannot forward messages on behalf of others. Each message is delivered directly sender $\to$ recipient.
  \item \textbf{No DHT/bootstrap}: Iroh does not perform peer discovery. Discovery happens via BLE ANNOUNCE packets or out-of-band exchange.
  \item \textbf{BLE is preferred}: When both transports are available, BLE is used first (lower latency, no Internet dependency).
\end{enumerate}

\subsection{Address Types Under Analysis}

\begin{table}[H]
\centering
\begin{tabular}{llp{7cm}}
\toprule
\textbf{Address Type} & \textbf{Abbreviation} & \textbf{Description} \\
\midrule
BLE only & BLE & Bluetooth Low Energy; $\sim$10--100m range, no IP stack \\
Private IPv4 & PIv4 & Behind NAT (e.g., \texttt{192.168.x.x}, \texttt{10.x.x.x}); not routable from Internet \\
Public IPv4 & PuIv4 & Directly routable IPv4; rare for mobile devices \\
Link-local IPv6 & LLv6 & \texttt{fe80::/10}; only valid on local link, not routable \\
Public IPv6 & PuIv6 & Globally routable IPv6; increasingly common on mobile carriers \\
\bottomrule
\end{tabular}
\caption{Network address types evaluated in this document.}
\end{table}

% =============================================================================
\section{Transport Mechanics}
% =============================================================================

\subsection{BLE Transport}

\begin{itemize}
  \item \textbf{Range}: Typically 10--30m indoors, up to 100m outdoors with line of sight.
  \item \textbf{Discovery}: Each device advertises a unique service UUID derived from its Ed25519 public key. Scanning discovers all nearby BLE devices; service discovery confirms Bitchat peers.
  \item \textbf{Data}: After GATT service discovery and ANNOUNCE exchange, data flows over the Bitchat GATT characteristic.
  \item \textbf{No IP required}: BLE operates below the IP layer. Works without any Internet or WiFi connectivity.
  \item \textbf{Symmetric}: Both peers must be in range; there is no multi-hop BLE.
\end{itemize}

\subsection{Iroh Transport}

\begin{itemize}
  \item \textbf{Protocol}: QUIC over UDP. Encrypted, multiplexed, and supports connection migration.
  \item \textbf{Identity}: NodeId = Ed25519 public key. The same key used for BLE identity.
  \item \textbf{Addressing}: A peer is addressed by \texttt{(NodeId, relay\_url, direct\_addresses[])}.
  \item \textbf{Relay servers}: Operated by iroh.computer (e.g., \texttt{https://use1-1.relay.iroh.network.}). They:
    \begin{itemize}
      \item Relay encrypted QUIC packets between peers who cannot reach each other directly
      \item Coordinate hole-punching attempts for NAT traversal
      \item Do NOT inspect, store, or modify message content
    \end{itemize}
  \item \textbf{Connection lifecycle}:
    \begin{enumerate}
      \item Peer A knows Peer B's NodeId + relay URL (from BLE ANNOUNCE or out-of-band)
      \item A connects to B via the relay server
      \item Iroh attempts hole-punching in the background
      \item If hole-punch succeeds, connection migrates to direct UDP (relay no longer used)
      \item If hole-punch fails, relay continues to carry traffic
    \end{enumerate}
  \item \textbf{QUIC connection migration}: When a device changes IP (e.g., WiFi $\to$ cellular), QUIC can migrate the connection to the new address without re-establishing.
\end{itemize}

% =============================================================================
\section{Peer Topology Definitions}
% =============================================================================

We evaluate four peer topologies:

\begin{enumerate}[label=\textbf{T\arabic*}]
  \item \textbf{Friend Next to Me} --- Both peers are physically co-located ($<$30m). Both BLE and iroh are potentially available.
  \item \textbf{Friend Far Away} --- Peers are geographically separated. Only iroh (Internet) is possible. No BLE.
  \item \textbf{Potential Relay Peer (``Friend in the Middle'')} --- A third peer C exists between peers A and B. We evaluate whether C can facilitate A$\leftrightarrow$B communication. \textit{Spoiler: under current design, it cannot.}
  \item \textbf{Server} --- Either (a) iroh's public relay infrastructure, or (b) a custom always-on supernode with a stable public IP.
\end{enumerate}

% =============================================================================
\section{Connectivity Matrix: Friend Next to Me (T1)}
% =============================================================================

Both peers are within BLE range ($<$30m). Each row represents the peer's network address type. We evaluate whether each transport works.

\begin{table}[H]
\centering
\small
\begin{tabular}{L{2.2cm} C{1.5cm} C{1.5cm} C{1.5cm} C{5cm}}
\toprule
\textbf{My Address} & \textbf{Peer's Address} & \textbf{BLE} & \textbf{Iroh} & \textbf{Notes} \\
\midrule
BLE only & BLE only & \yes & \no & BLE works. No IP = no iroh. \\
\rowcolor{lightgray}
BLE only & PIv4 & \yes & \no & I have no IP; cannot reach relay or peer. \\
BLE only & PuIv4 & \yes & \no & Same---I have no IP stack for iroh. \\
\rowcolor{lightgray}
BLE only & LLv6 & \yes & \no & Link-local IPv6 is not routable to relay. \\
BLE only & PuIv6 & \yes & \no & I have no IP at all. \\
\midrule
PIv4 & PIv4 & \yes & \prt & Both behind NAT. Relay-mediated; hole-punch may succeed if same LAN. \\
\rowcolor{lightgray}
PIv4 & PuIv4 & \yes & \yes & I reach relay; peer is directly reachable. Hole-punch succeeds. \\
PIv4 & LLv6 & \yes & \no & Link-local not routable. Peer cannot reach relay. \\
\rowcolor{lightgray}
PIv4 & PuIv6 & \yes & \yes & Cross-protocol via relay. Direct only if I also have IPv6. \\
\midrule
PuIv4 & PuIv4 & \yes & \yes & Both public. Direct connection, no relay needed. \\
\rowcolor{lightgray}
PuIv4 & PIv4 & \yes & \yes & I'm reachable; peer connects via relay then hole-punches to me. \\
PuIv4 & LLv6 & \yes & \no & Peer cannot reach relay or me. \\
\rowcolor{lightgray}
PuIv4 & PuIv6 & \yes & \yes & Direct if both stacks available; relay bridges otherwise. \\
\midrule
LLv6 & LLv6 & \yes & \prt & Same link: iroh \textit{might} work if UDP on link-local. Unreliable. \\
\rowcolor{lightgray}
LLv6 & Any public & \yes & \no & I cannot reach relay. \\
\midrule
PuIv6 & PuIv6 & \yes & \yes & Direct IPv6 connection. No relay needed. \\
\rowcolor{lightgray}
PuIv6 & PIv4 & \yes & \yes & Via relay. Hole-punch unlikely (different protocols). \\
PuIv6 & PuIv4 & \yes & \yes & Direct possible if peer's stack supports IPv6 too. \\
\bottomrule
\end{tabular}
\caption{Connectivity matrix for co-located peers (T1). BLE always works regardless of IP.}
\end{table}

\subsection{T1 Analysis}

\textbf{Key insight}: When peers are co-located, \textbf{BLE is always available} regardless of network configuration. This makes the IP address type irrelevant for basic connectivity---BLE provides the reliable floor.

Iroh adds value even when co-located:
\begin{itemize}
  \item Higher throughput than BLE (QUIC vs GATT characteristic writes)
  \item Can carry larger payloads without fragmentation
  \item But requires both peers to have routable IP addresses (or relay access)
\end{itemize}

\textbf{Worst case}: Both peers have only BLE (no IP). Communication works but is limited to BLE throughput ($\sim$1--2 KB/s effective after protocol overhead).

\textbf{Best case}: Both peers have public IPs. BLE handles discovery + low-latency small messages; iroh handles bulk data transfer with direct QUIC connection.

% =============================================================================
\section{Connectivity Matrix: Friend Far Away (T2)}
% =============================================================================

Peers are geographically separated. BLE is not available. Only iroh can work.

\begin{table}[H]
\centering
\small
\begin{tabular}{L{2.2cm} C{1.8cm} C{1.3cm} C{6cm}}
\toprule
\textbf{My Address} & \textbf{Peer's Address} & \textbf{Iroh} & \textbf{Notes} \\
\midrule
BLE only & Any & \no & No IP = no iroh. \textbf{Unreachable.} \\
\rowcolor{lightgray}
PIv4 & BLE only & \no & Peer has no IP. \textbf{Unreachable.} \\
PIv4 & PIv4 & \dgr & Both behind NAT. Relay works. Hole-punch success depends on NAT type (see \S\ref{sec:nat}). \\
\rowcolor{lightgray}
PIv4 & PuIv4 & \yes & I connect via relay; hole-punch to peer's public IP likely succeeds. \\
PIv4 & LLv6 & \no & Peer's address not routable. \\
\rowcolor{lightgray}
PIv4 & PuIv6 & \dgr & Relay bridges IPv4$\leftrightarrow$IPv6. Direct unlikely without dual-stack. \\
\midrule
PuIv4 & PIv4 & \yes & Peer connects to me via relay; hole-punch succeeds (I'm public). \\
\rowcolor{lightgray}
PuIv4 & PuIv4 & \yes & Direct connection. No relay needed. \\
PuIv4 & PuIv6 & \yes & Relay bridges protocols. Direct if both dual-stack. \\
\midrule
\rowcolor{lightgray}
LLv6 & Any & \no & Cannot reach relay or any remote peer. \textbf{Unreachable.} \\
\midrule
PuIv6 & PIv4 & \dgr & Relay bridges. Hole-punch unlikely cross-protocol. \\
\rowcolor{lightgray}
PuIv6 & PuIv4 & \yes & Relay or direct if both dual-stack. \\
PuIv6 & PuIv6 & \yes & Direct IPv6 connection. Optimal. \\
\bottomrule
\end{tabular}
\caption{Connectivity matrix for remote peers (T2). No BLE available.}
\end{table}

\subsection{T2 Analysis}

\textbf{Critical dependency}: Without BLE, the \textit{only} transport is iroh. If either peer lacks a routable IP address (BLE-only or link-local IPv6), communication is \textbf{impossible}.

\subsubsection{NAT Traversal Success Rates}
\label{sec:nat}

Iroh's hole-punching success depends on NAT types:

\begin{table}[H]
\centering
\begin{tabular}{lcccc}
\toprule
& \textbf{Full Cone} & \textbf{Restricted} & \textbf{Port Restricted} & \textbf{Symmetric} \\
\midrule
\textbf{Full Cone} & \yes & \yes & \yes & \yes \\
\textbf{Restricted} & \yes & \yes & \yes & \prt \\
\textbf{Port Restricted} & \yes & \yes & \prt & \no \\
\textbf{Symmetric} & \yes & \prt & \no & \no \\
\bottomrule
\end{tabular}
\caption{Hole-punching success by NAT type combination. Symmetric NAT (common on carrier-grade NAT / mobile networks) is the hardest case.}
\end{table}

When hole-punching fails, iroh falls back to relay. Relay always works as long as both peers can reach the relay server over TCP/HTTPS. This means:

\begin{itemize}
  \item \textbf{PIv4 $\leftrightarrow$ PIv4}: Always works via relay even if hole-punch fails
  \item \textbf{Symmetric NAT $\leftrightarrow$ Symmetric NAT}: Relay-only, higher latency ($\sim$50--150ms added round-trip)
\end{itemize}

% =============================================================================
\section{Connectivity Matrix: Potential Relay Peer (T3)}
% =============================================================================

\textbf{Scenario}: Peer A wants to reach Peer B. Peer C is a third peer connected to both A and B. Can C help?

\begin{table}[H]
\centering
\begin{tabular}{L{3cm} C{2cm} C{6cm}}
\toprule
\textbf{Scenario} & \textbf{A$\to$B via C?} & \textbf{Reason} \\
\midrule
A$\leftrightarrow$C (BLE), C$\leftrightarrow$B (BLE) & \no & Bitchat does not relay. C cannot forward A's messages to B. \\
\rowcolor{lightgray}
A$\leftrightarrow$C (BLE), C$\leftrightarrow$B (iroh) & \no & Same. No store-and-forward, no relaying. \\
A$\leftrightarrow$C (iroh), C$\leftrightarrow$B (iroh) & \no & C is not a relay server. Bitchat protocol forbids relaying. \\
\rowcolor{lightgray}
A$\leftrightarrow$C (BLE), C$\leftrightarrow$B (BLE), B out of A's range & \no & No BLE mesh / multi-hop. \\
Any combination & \no & \textbf{By design}: no intermediate peer can relay messages. \\
\bottomrule
\end{tabular}
\caption{Relay peer analysis (T3). Result: always \textcolor{no}{\textbf{No}}.}
\end{table}

\subsection{T3 Analysis}

This is a deliberate architectural decision:

\begin{itemize}
  \item \textbf{Simplicity}: No routing tables, no hop counts, no TTLs, no duplicate detection.
  \item \textbf{Privacy}: Messages never traverse third-party peers. Only iroh relay servers (which see only encrypted QUIC packets) sit in the path.
  \item \textbf{Reliability}: No dependency on intermediate peers staying online.
  \item \textbf{Trade-off}: If A and B are both within BLE range of C but not of each other, they simply cannot communicate. This is accepted.
\end{itemize}

\subsubsection{What Would Change If Relaying Were Added}

For reference only (this is \textbf{NOT} planned):

\begin{itemize}
  \item BLE mesh would extend effective range: A$\to$C$\to$B adds $\sim$60m effective range per hop
  \item But: latency doubles per hop, throughput halves, battery drain increases significantly
  \item Routing becomes a hard problem (which peer to relay through? What if C moves?)
  \item Message duplication and loops must be handled
  \item Privacy model changes entirely---C sees message metadata
\end{itemize}

% =============================================================================
\section{Server Scenarios (T4)}
% =============================================================================

\subsection{T4a: Iroh Public Relay Infrastructure}

Iroh provides free relay servers. These are the \textbf{current default}.

\begin{table}[H]
\centering
\small
\begin{tabular}{L{3cm} C{1.5cm} C{7cm}}
\toprule
\textbf{Scenario} & \textbf{Works?} & \textbf{Details} \\
\midrule
Both peers behind NAT & \yes & Relay carries traffic. Hole-punch attempted. \\
\rowcolor{lightgray}
One peer public IP & \yes & Relay + hole-punch. Migrates to direct quickly. \\
Both peers public IP & \yes & Can connect directly; relay not actually used. \\
\rowcolor{lightgray}
Relay server down & \dgr & Only direct connections work. PIv4$\leftrightarrow$PIv4 fails entirely. \\
Both peers BLE only & \no & No IP connectivity at all. \\
\rowcolor{lightgray}
Cross IPv4/IPv6 & \yes & Relay bridges protocol families. \\
\bottomrule
\end{tabular}
\caption{Iroh public relay infrastructure scenarios.}
\end{table}

\textbf{Risks of relying on iroh's public relays}:
\begin{itemize}
  \item \textbf{Availability}: If iroh.computer's relay infrastructure goes down, NAT traversal breaks. Peers with public IPs are unaffected.
  \item \textbf{Censorship}: Relay domains could be blocked in certain jurisdictions.
  \item \textbf{Rate limits}: Public relays may impose bandwidth or connection limits.
  \item \textbf{Latency}: Geographic distance to nearest relay adds latency before hole-punch succeeds.
\end{itemize}

\subsection{T4b: Custom Always-On Supernode}

A server with a stable public IP running a Bitchat peer (or iroh relay). Characteristics: always online, public IP, high bandwidth.

\begin{table}[H]
\centering
\small
\begin{tabular}{L{3.5cm} C{1.5cm} C{6.5cm}}
\toprule
\textbf{Scenario} & \textbf{Works?} & \textbf{Details} \\
\midrule
As custom relay server & \yes & Peers configure its URL as relay. Full control over availability. \\
\rowcolor{lightgray}
As always-on peer & \yes & Can exchange messages with any peer that can reach it. Always discoverable. \\
As message relay for other peers & \no & Bitchat forbids relaying. Server cannot forward A's messages to B. \\
\rowcolor{lightgray}
As discovery hub & \prt & Could distribute peer address lists via iroh, but currently ANNOUNCE is BLE-only for discovery. Would need protocol extension. \\
As fallback when public relay is down & \yes & Self-hosted relay provides redundancy. \\
\bottomrule
\end{tabular}
\caption{Custom supernode scenarios.}
\end{table}

\textbf{Value proposition of a supernode}:
\begin{itemize}
  \item \textbf{Relay redundancy}: If iroh's public relays go down, your server still works.
  \item \textbf{Always-reachable peer}: Useful for peers who need a ``mailbox'' (but remember: no store-and-forward, so both must be online simultaneously).
  \item \textbf{Discovery service}: Could extend ANNOUNCE beyond BLE range by acting as an address book. \textit{Not currently implemented.}
\end{itemize}

\textbf{What a supernode does NOT solve}:
\begin{itemize}
  \item Offline message delivery (no store-and-forward by design)
  \item BLE-only peers still cannot reach it
  \item Cannot relay messages between other peers
\end{itemize}

% =============================================================================
\section{Mobility Analysis}
% =============================================================================

\subsection{M1: WiFi to Cellular Handoff}

\begin{tikzpicture}[
  node distance=2.5cm,
  peer/.style={rectangle, draw, rounded corners, minimum width=2cm, minimum height=1cm, fill=blue!10},
  event/.style={rectangle, draw, dashed, rounded corners, minimum width=3cm, fill=yellow!10},
  arrow/.style={-{Stealth[length=3mm]}, thick}
]
  \node[peer] (A) {Peer A};
  \node[event, right=of A] (handoff) {WiFi $\to$ LTE};
  \node[peer, right=of handoff] (B) {Peer B};

  \draw[arrow, color=yes] (A) -- node[above, font=\small] {QUIC} (handoff);
  \draw[arrow, color=yes] (handoff) -- node[above, font=\small] {migrates} (B);
  \draw[arrow, color=degraded, bend right=30] (A) to node[below, font=\small] {brief interruption} (B);
\end{tikzpicture}

\begin{table}[H]
\centering
\small
\begin{tabular}{L{3cm} L{2cm} L{2cm} L{5cm}}
\toprule
\textbf{Before} & \textbf{After} & \textbf{Result} & \textbf{Details} \\
\midrule
WiFi (PIv4) & LTE (PIv4) & \dgr & IP changes. QUIC connection migration kicks in. Brief interruption ($<$5s typical). Relay reconnection needed. \\
\rowcolor{lightgray}
WiFi (PIv4) & LTE (PuIv6) & \dgr & Protocol family change. Relay rebinds. Connection survives via QUIC migration if relay supports both. \\
WiFi (PuIv6) & LTE (PIv4) & \dgr & Downgrade from public to NATed. Direct connection lost; falls back to relay. \\
\rowcolor{lightgray}
WiFi + BLE & LTE (no BLE peer) & \prt & BLE connection lost (peer out of range or WiFi peer only). Iroh takes over if available. \\
\bottomrule
\end{tabular}
\caption{WiFi to cellular handoff scenarios.}
\end{table}

\textbf{Analysis}: QUIC's connection migration is the key enabler here. Unlike TCP, QUIC connections are identified by a connection ID, not by the 4-tuple (src IP, src port, dst IP, dst port). When the device's IP changes, iroh detects the change and re-establishes the path through the relay. The connection survives with a brief interruption.

\textbf{BLE impact}: If the peer was also a nearby BLE peer, the BLE connection is unaffected by the IP change. BLE and iroh are independent transports.

\subsection{M2: Moving Between WiFi Networks}

\begin{table}[H]
\centering
\small
\begin{tabular}{L{3cm} L{2cm} L{2cm} L{5cm}}
\toprule
\textbf{Before} & \textbf{After} & \textbf{Result} & \textbf{Details} \\
\midrule
WiFi A (PIv4) & WiFi B (PIv4) & \dgr & New private IP. QUIC migrates via relay. $\sim$2--10s interruption during DHCP + reconnect. \\
\rowcolor{lightgray}
WiFi A (same LAN as peer) & WiFi B (different LAN) & \dgr & Direct LAN connection lost. Falls back to relay. Hole-punch may re-establish direct. \\
Same AP as peer & Different AP, same network & \yes & Private IP may not change. BLE may drop if range exceeded. \\
\rowcolor{lightgray}
WiFi A & Brief disconnect & \dgr & QUIC has idle timeout. If reconnect within timeout, connection survives. Otherwise, full reconnection needed. \\
\bottomrule
\end{tabular}
\caption{WiFi-to-WiFi transition scenarios.}
\end{table}

\textbf{Gap period}: Between disconnecting from WiFi A and getting a DHCP lease on WiFi B, there is a complete connectivity gap (typically 2--10 seconds). During this gap:
\begin{itemize}
  \item Iroh connections are suspended (no packets flow)
  \item BLE continues to work if peer is in range
  \item Messages sent during the gap fail immediately (no store-and-forward)
\end{itemize}

\subsection{M3: BLE Range Transitions}

\begin{table}[H]
\centering
\small
\begin{tabular}{L{3.5cm} L{1.5cm} L{1.5cm} L{5cm}}
\toprule
\textbf{Scenario} & \textbf{BLE} & \textbf{Iroh} & \textbf{Details} \\
\midrule
Peer walks into BLE range & \yes & unchanged & BLE scan discovers peer. ANNOUNCE exchanged. New transport available. \\
\rowcolor{lightgray}
Peer walks out of BLE range & \no & unchanged & BLE disconnects (detected via RSSI drop / timeout). Iroh unaffected if both have IP. \\
Peer walks out, no Internet & \no & \no & Both transports lost. \textbf{Peer unreachable.} \\
\rowcolor{lightgray}
Peer walks in, both have iroh & \yes & \yes & BLE added as preferred transport. Iroh remains as fallback. \\
BLE flapping (edge of range) & \dgr & unchanged & Repeated connect/disconnect. Should debounce in BLE layer. Iroh provides stable fallback. \\
\bottomrule
\end{tabular}
\caption{BLE range transition scenarios.}
\end{table}

\textbf{Transport failover}: When BLE drops and iroh is available, messages automatically route via iroh. When BLE becomes available again, it is preferred. This failover is transparent to the application layer.

\subsection{M4: BLE-Only Operation (No Internet)}

This is the most constrained scenario: the device has no IP connectivity at all.

\begin{table}[H]
\centering
\small
\begin{tabular}{L{4cm} C{1.5cm} C{6cm}}
\toprule
\textbf{Scenario} & \textbf{Works?} & \textbf{Details} \\
\midrule
Peer next to me, both BLE-only & \yes & Full communication via BLE. Limited to BLE throughput. \\
\rowcolor{lightgray}
Peer 50m away, both BLE-only & \prt & At BLE range limit. Unreliable, frequent disconnects. \\
Peer 200m away, both BLE-only & \no & Out of BLE range. No alternative transport. \\
\rowcolor{lightgray}
3 peers in a line, BLE-only & \no & Middle peer in range of both, but no relaying. A$\leftrightarrow$C only if in direct range. \\
Airplane mode with BLE on & \yes & BLE works independently of airplane mode radio restrictions (on most devices). \\
\rowcolor{lightgray}
Underground / no cell signal & \yes & BLE unaffected. Only local peers reachable. \\
\bottomrule
\end{tabular}
\caption{BLE-only operation scenarios (no Internet).}
\end{table}

\textbf{BLE-only implications}:
\begin{itemize}
  \item \textbf{Range is the hard limit}: $\sim$10--30m indoors, $\sim$100m outdoors. No way to extend without relaying (which is forbidden).
  \item \textbf{Throughput}: BLE GATT writes are $\sim$1--2 KB/s effective. Fine for text messages, insufficient for media.
  \item \textbf{Discovery}: BLE scanning is the only discovery mechanism. No Internet means no out-of-band discovery.
  \item \textbf{Resilient}: Works in disaster scenarios, underground, in Faraday cages, on planes---anywhere Bluetooth radio works.
\end{itemize}

% =============================================================================
\section{Combined Scenario Matrix}
% =============================================================================

This master matrix combines address type, topology, and mobility for the most common real-world scenarios.

\begin{longtable}{L{3.5cm} L{2.5cm} C{1cm} C{1cm} L{3.5cm}}
\toprule
\textbf{Scenario} & \textbf{Address Config} & \textbf{BLE} & \textbf{Iroh} & \textbf{Effective Connectivity} \\
\midrule
\endfirsthead
\toprule
\textbf{Scenario} & \textbf{Address Config} & \textbf{BLE} & \textbf{Iroh} & \textbf{Effective Connectivity} \\
\midrule
\endhead
\midrule
\multicolumn{5}{r}{\textit{continued on next page}} \\
\bottomrule
\endfoot
\bottomrule
\endlastfoot

% --- Co-located scenarios ---
\multicolumn{5}{l}{\textbf{Co-located Peers ($<$30m)}} \\
\midrule
Coffee shop, both on WiFi & PIv4 / PIv4 & \yes & \yes & Excellent. BLE preferred, iroh fallback. Likely same LAN = direct iroh. \\
\rowcolor{lightgray}
Coffee shop, one on LTE & PIv4 / PIv4 & \yes & \dgr & BLE works. Iroh via relay (different NATs). \\
Park, no WiFi & BLE / LTE(PIv4) & \yes & \prt & BLE works. One peer has iroh. \\
\rowcolor{lightgray}
Underground, no signal & BLE / BLE & \yes & \no & BLE only. Works but limited. \\

\midrule
% --- Remote scenarios ---
\multicolumn{5}{l}{\textbf{Remote Peers (different cities)}} \\
\midrule
Both on home WiFi & PIv4 / PIv4 & \no & \dgr & Relay only. Hole-punch depends on router NAT type. \\
\rowcolor{lightgray}
One on fiber (PuIv4) & PuIv4 / PIv4 & \no & \yes & Direct connection to public peer. Fast. \\
Both on LTE & PIv4 / PIv4 & \no & \dgr & Carrier-grade NAT = symmetric. Relay-only likely. \\
\rowcolor{lightgray}
Both have IPv6 & PuIv6 / PuIv6 & \no & \yes & Direct IPv6. Optimal for remote. \\
One offline & Any / BLE only & \no & \no & \textbf{Unreachable.} \\

\midrule
% --- Mobility scenarios ---
\multicolumn{5}{l}{\textbf{Mobility}} \\
\midrule
Walking away from BLE peer, have LTE & PIv4 & $\to$\no & \yes & BLE drops. Iroh takes over seamlessly. \\
\rowcolor{lightgray}
Walking toward peer, gaining BLE & PIv4 & $\to$\yes & \yes & BLE discovered. Becomes preferred transport. \\
On train, passing through cells & PIv4 $\to$ PIv4 & \no & \dgr & Frequent IP changes. QUIC migration helps. Intermittent drops. \\
\rowcolor{lightgray}
Entering airplane mode + BLE & BLE & \yes & $\to$\no & Iroh lost. BLE works for nearby peers. \\
Leaving subway (no signal $\to$ LTE) & BLE $\to$ PIv4 & \yes & $\to$\yes & Iroh becomes available. Remote peers now reachable. \\
\rowcolor{lightgray}
WiFi drops, no cellular & PIv4 $\to$ BLE & \yes & $\to$\no & Only nearby BLE peers reachable. \\

\end{longtable}

% =============================================================================
\section{Gap Analysis and Recommendations}
% =============================================================================

\subsection{Identified Gaps}

\begin{enumerate}
  \item \textbf{BLE-only peers cannot reach remote peers}: This is by design but limits utility in disconnected scenarios. No mitigation without adding relaying (rejected by design).

  \item \textbf{Discovery requires BLE or out-of-band}: If two peers have never been co-located and have no out-of-band channel, they cannot discover each other's iroh addresses. A discovery service could address this.

  \item \textbf{Symmetric NAT + Symmetric NAT}: Common on mobile networks. Always relay-only. Latency impact: $+$50--150ms RTT. Mitigation: iroh relay servers are geographically distributed.

  \item \textbf{IPv4/IPv6 mismatch}: When one peer has only IPv4 and the other only IPv6, direct connection is impossible. Relay bridges this but adds latency.

  \item \textbf{Relay server dependency}: All NAT-traversal scenarios depend on iroh's relay infrastructure. If it goes down, only direct-IP and BLE work.

  \item \textbf{No offline delivery}: If the recipient is offline, the message fails. The sender must manually retry. No notification when the peer comes back online (unless they re-enter BLE range).
\end{enumerate}

\subsection{Recommendations}

\begin{table}[H]
\centering
\begin{tabular}{C{0.5cm} L{4cm} L{3cm} L{4cm}}
\toprule
\textbf{\#} & \textbf{Gap} & \textbf{Recommendation} & \textbf{Impact} \\
\midrule
1 & Relay server dependency & Support custom relay URLs in config & Enables self-hosted redundancy \\
\rowcolor{lightgray}
2 & Discovery beyond BLE & Address exchange via iroh (QR code, link) & Enables remote-first friendships \\
3 & Symmetric NAT latency & Multiple relay regions & Reduce relay RTT to $<$50ms \\
\rowcolor{lightgray}
4 & No peer-online notification & Lightweight presence ping via iroh & Know when a remote peer is reachable \\
5 & BLE range limitation & Accept as design constraint & Simplicity over range \\
\bottomrule
\end{tabular}
\caption{Recommendations for identified connectivity gaps.}
\end{table}

% =============================================================================
\section{Historical Comparison: LibP2P (Previous Stack)}
% =============================================================================

Before adopting iroh, Bitchat used \textbf{libp2p} (specifically \texttt{dart\_libp2p} with \texttt{dart\_udx}) as its Internet transport layer. This section reconstructs the connectivity assumptions under libp2p and compares them with the current iroh-based architecture.

\subsection{LibP2P Transport Mechanics}

\begin{itemize}
  \item \textbf{Transports}: Two transport protocols were available:
    \begin{itemize}
      \item \textbf{UDX} (UDP-based, custom protocol) --- primary transport, preferred for performance
      \item \textbf{TCP} --- secondary, used for IPFS ecosystem interoperability
    \end{itemize}
  \item \textbf{Security}: Noise protocol for all connections (both UDX and TCP).
  \item \textbf{Identity}: LibP2P used its own Ed25519 keypair to generate a \texttt{PeerId}, \textbf{separate} from the Bitchat application-layer Ed25519 public key. This required maintaining a double mapping:
    \begin{enumerate}
      \item BLE Device ID $\leftrightarrow$ LibP2P PeerId
      \item LibP2P PeerId $\leftrightarrow$ Bitchat Public Key
    \end{enumerate}
  \item \textbf{Addressing}: Multiaddr format (e.g., \texttt{/ip4/1.2.3.4/tcp/4001/p2p/12D3Koo...})
  \item \textbf{Application protocol}: \texttt{/gever/1.0.0} --- each message sent via a new stream (not persistent connection)
  \item \textbf{DHT}: Supported but \textbf{disabled} (\texttt{enableDht = false}). Bitchat never used Kademlia for peer discovery.
  \item \textbf{mDNS}: \textbf{Enabled} (\texttt{enableMdns = true}). Provided automatic local network peer discovery.
\end{itemize}

\subsection{LibP2P NAT Traversal Stack}

LibP2P provided a multi-layered NAT traversal strategy:

\begin{enumerate}
  \item \textbf{AutoNAT v2}: Probes reachability by asking other peers to dial back. Classifies the local NAT type (public, restricted, symmetric).
  \item \textbf{Circuit Relay v2}: Public peers volunteer as relays. NATed peers reserve relay slots. Traffic flows: A $\to$ Relay Peer $\to$ B. \textit{Unlike iroh relays, these are libp2p peers themselves, not dedicated infrastructure.}
  \item \textbf{AutoRelay}: Automatically discovers and connects to available relay peers when AutoNAT detects the node is behind NAT.
  \item \textbf{DCUtR (Direct Connection Upgrade through Relay)}: After establishing a relayed connection, peers coordinate hole-punching to upgrade to a direct connection.
\end{enumerate}

\textbf{Address priority hierarchy} (highest to lowest):
\begin{enumerate}
  \item IPv6/UDX --- best performance, globally routable
  \item IPv4/UDX --- common, good performance
  \item IPv4/TCP --- IPFS interop, wider firewall compatibility
  \item Circuit Relay --- last resort, highest latency
\end{enumerate}

\subsection{LibP2P Bootstrap Requirement}

LibP2P \textbf{required bootstrap peers} for relay discovery:

\begin{verbatim}
/ip4/40.160.9.115/tcp/4001/p2p/12D3KooWKnDdG3...
/ip6/2604:2dc0:101:100::138f/tcp/4001/p2p/12D3KooWKnDdG3...
\end{verbatim}

These bootstrap peers served as:
\begin{itemize}
  \item Initial connection points for newly started nodes
  \item Relay reservation targets (via Circuit Relay v2)
  \item AutoNAT probing partners
\end{itemize}

Without the bootstrap peer being reachable, a NATed node could not discover relays and was limited to direct connections or mDNS.

\subsection{Connectivity Matrix: Friend Next to Me (LibP2P)}

\begin{table}[H]
\centering
\small
\begin{tabular}{L{2.2cm} C{1.5cm} C{1.3cm} C{1.3cm} C{4.5cm}}
\toprule
\textbf{My Addr} & \textbf{Peer Addr} & \textbf{BLE} & \textbf{LP2P} & \textbf{Notes} \\
\midrule
BLE only & BLE only & \yes & \no & No IP. Identical to iroh. \\
\rowcolor{lightgray}
BLE only & PIv4 & \yes & \no & I have no IP. Identical to iroh. \\
BLE only & PuIv4 & \yes & \no & Same. \\
\rowcolor{lightgray}
BLE only & LLv6 & \yes & \no & Same. \\
BLE only & PuIv6 & \yes & \no & Same. \\
\midrule
PIv4 & PIv4 & \yes & \yes$^*$ & \textbf{Differs from iroh}: mDNS discovers peer on same LAN without relay. If different LANs: needs bootstrap $\to$ relay. \\
\rowcolor{lightgray}
PIv4 & PuIv4 & \yes & \yes & Direct TCP or UDX to public peer. No relay needed. \\
PIv4 & LLv6 & \yes & \prt & mDNS might discover on same link. UDX over link-local unreliable. \\
\rowcolor{lightgray}
PIv4 & PuIv6 & \yes & \dgr & Cross-protocol. Needs relay peer that bridges IPv4$\leftrightarrow$IPv6. Not guaranteed. \\
\midrule
PuIv4 & PuIv4 & \yes & \yes & Direct. No relay needed. \\
\rowcolor{lightgray}
PuIv4 & PIv4 & \yes & \yes & Peer connects to me directly or via relay. \\
PuIv4 & LLv6 & \yes & \no & Link-local not routable. mDNS possible on same link only. \\
\rowcolor{lightgray}
PuIv4 & PuIv6 & \yes & \dgr & Cross-protocol relay uncertain. Relay peer must be dual-stack. \\
\midrule
LLv6 & LLv6 & \yes & \prt & mDNS on same link. UDX \textit{might} bind link-local. Better than iroh here due to mDNS. \\
\rowcolor{lightgray}
LLv6 & Any public & \yes & \no & Cannot reach bootstrap or peer. \\
\midrule
PuIv6 & PuIv6 & \yes & \yes & Direct IPv6/UDX. Optimal. \\
\rowcolor{lightgray}
PuIv6 & PIv4 & \yes & \dgr & Cross-protocol via relay. Relay must be dual-stack. \\
PuIv6 & PuIv4 & \yes & \dgr & Same cross-protocol issue. \\
\bottomrule
\end{tabular}
\caption{LibP2P co-located connectivity. $^*$mDNS provides same-LAN discovery that iroh lacks.}
\end{table}

\subsection{Connectivity Matrix: Friend Far Away (LibP2P)}

\begin{table}[H]
\centering
\small
\begin{tabular}{L{2.2cm} C{1.8cm} C{1.3cm} C{5.5cm}}
\toprule
\textbf{My Addr} & \textbf{Peer Addr} & \textbf{LP2P} & \textbf{Notes} \\
\midrule
BLE only & Any & \no & No IP. Identical to iroh. \\
\rowcolor{lightgray}
PIv4 & BLE only & \no & Peer has no IP. \\
PIv4 & PIv4 & \dgr & Both NATed. Need bootstrap peer reachable. AutoRelay finds relay. DCUtR attempts hole-punch. More moving parts than iroh relay. \\
\rowcolor{lightgray}
PIv4 & PuIv4 & \yes & Direct to peer. Can use TCP or UDX. \\
PIv4 & LLv6 & \no & Not routable. \\
\rowcolor{lightgray}
PIv4 & PuIv6 & \dgr & Need dual-stack relay peer. Not guaranteed in libp2p relay pool. \\
\midrule
PuIv4 & PIv4 & \yes & Peer uses relay then DCUtR to me. \\
\rowcolor{lightgray}
PuIv4 & PuIv4 & \yes & Direct. Both UDX and TCP work. \\
PuIv4 & PuIv6 & \dgr & Cross-protocol. Relay or direct only if both dual-stack. \\
\midrule
\rowcolor{lightgray}
LLv6 & Any & \no & Cannot reach bootstrap or any remote peer. \\
\midrule
PuIv6 & PIv4 & \dgr & Cross-protocol relay required. \\
\rowcolor{lightgray}
PuIv6 & PuIv4 & \dgr & Same issue. Direct only if peer also has IPv6. \\
PuIv6 & PuIv6 & \yes & Direct IPv6/UDX. Best case. \\
\bottomrule
\end{tabular}
\caption{LibP2P remote connectivity. Cross-protocol scenarios are worse than iroh due to relay peer heterogeneity.}
\end{table}

\subsection{Potential Relay Peer Under LibP2P}

This is where libp2p and iroh \textbf{differ fundamentally}.

\begin{table}[H]
\centering
\small
\begin{tabular}{L{4cm} C{1.5cm} C{1.5cm} C{4.5cm}}
\toprule
\textbf{Scenario} & \textbf{LP2P} & \textbf{Iroh} & \textbf{Details} \\
\midrule
C is a public peer, A and B are NATed & \yes$^*$ & \no & LibP2P: C can serve as Circuit Relay v2 if it has relay capability enabled. A and B reserve slots on C. \textit{But}: Bitchat application layer still forbids message relaying; this is transport-level only. \\
\rowcolor{lightgray}
C is NATed too & \no & \no & NATed peers cannot serve as relays in either stack. \\
C has BLE to A, iroh to B & \no & \no & Even with libp2p, the BLE$\to$IP bridge was not implemented. \\
\rowcolor{lightgray}
C is the bootstrap peer & \yes & \na & Bootstrap peer doubles as relay. This was the primary relay path. \\
\bottomrule
\end{tabular}
\caption{Relay peer comparison. $^*$LibP2P supported transport-level relay via Circuit Relay v2, but Bitchat never relayed application messages through peers.}
\end{table}

\textbf{Critical distinction}: LibP2P's Circuit Relay v2 allowed \textit{any public peer} to act as a transport-level relay. This meant:
\begin{itemize}
  \item A third peer C with a public IP could transparently relay encrypted libp2p connections between A and B
  \item This happened at the transport layer---C never saw application-level messages (Noise-encrypted)
  \item In practice, Bitchat primarily relied on the bootstrap peer as the relay, not arbitrary third peers
\end{itemize}

With iroh, relay is provided by \textbf{dedicated relay servers} (iroh.computer infrastructure), not by peers. No Bitchat peer ever relays for another peer at any layer.

\subsection{Server Scenarios Under LibP2P}

\subsubsection{Bootstrap Peer (LibP2P's ``Server'')}

The bootstrap peer was a server at \texttt{40.160.9.115} with both IPv4 and IPv6 addresses.

\begin{table}[H]
\centering
\small
\begin{tabular}{L{3.5cm} C{1.5cm} C{6cm}}
\toprule
\textbf{Role} & \textbf{Works?} & \textbf{Details} \\
\midrule
Initial peer discovery & \yes & Newly started nodes connect to bootstrap to join the network. \\
\rowcolor{lightgray}
Relay server (Circuit Relay) & \yes & NATed peers reserve relay slots on bootstrap. Primary relay path. \\
AutoNAT probe target & \yes & Bootstrap dials back to test reachability. \\
\rowcolor{lightgray}
Single point of failure & \textcolor{no}{\textbf{Risk}} & If bootstrap goes down: no relay for NATed peers, no AutoNAT, new peers cannot join. \\
\bottomrule
\end{tabular}
\caption{Bootstrap peer roles under libp2p.}
\end{table}

\textbf{Single point of failure}: The bootstrap peer was a critical dependency. If it went offline:
\begin{itemize}
  \item Already-connected peers with direct connections continued working
  \item NATed peers lost their relay path (no new relay reservations)
  \item New peers could not join the network at all
  \item AutoNAT could not classify NAT type
\end{itemize}

\subsubsection{Custom Supernode (LibP2P vs Iroh)}

\begin{table}[H]
\centering
\small
\begin{tabular}{L{3cm} C{2cm} C{2cm} C{4.5cm}}
\toprule
\textbf{Capability} & \textbf{LP2P} & \textbf{Iroh} & \textbf{Comparison} \\
\midrule
Transport relay & \yes & \yes & LP2P: peer-level relay. Iroh: dedicated relay server. \\
\rowcolor{lightgray}
Always-on peer & \yes & \yes & Both support this equally. \\
Redundant relay & Hard & Easy & LP2P: adding relay peers requires config. Iroh: just add relay URL. \\
\rowcolor{lightgray}
Discovery hub & \prt & \prt & Both need protocol extensions for non-BLE discovery. \\
mDNS discovery & \yes & \no & LP2P supernode discoverable on LAN via mDNS. \\
\bottomrule
\end{tabular}
\caption{Supernode comparison between stacks.}
\end{table}

\subsection{Mobility Under LibP2P}

\subsubsection{WiFi to Cellular Handoff}

\begin{table}[H]
\centering
\small
\begin{tabular}{L{3cm} L{2cm} C{1.5cm} C{1.5cm} L{3.5cm}}
\toprule
\textbf{Transition} & \textbf{Transport} & \textbf{LP2P} & \textbf{Iroh} & \textbf{LP2P Details} \\
\midrule
WiFi (PIv4) $\to$ LTE (PIv4) & UDX/TCP & \no$^*$ & \dgr & \textbf{Connection dies}. UDX/TCP are bound to 4-tuple. No connection migration. Must fully reconnect: new connection to bootstrap, new relay reservation, re-establish peer connections. Recovery: 10--30s. \\
\rowcolor{lightgray}
WiFi (PIv4) $\to$ LTE (PuIv6) & UDX/TCP & \no$^*$ & \dgr & Worse: protocol family also changes. Full reconnection required. \\
WiFi + BLE $\to$ LTE & BLE & \yes & \yes & BLE unaffected (same as iroh). \\
\bottomrule
\end{tabular}
\caption{WiFi$\to$cellular handoff. $^*$LibP2P's UDX and TCP lack connection migration---this is where iroh's QUIC is significantly superior.}
\end{table}

\textbf{Key difference}: LibP2P over UDX/TCP \textbf{does not support connection migration}. When the IP changes:
\begin{enumerate}
  \item All existing connections (UDX and TCP) are severed
  \item The node must reconnect to the bootstrap peer
  \item AutoNAT must re-probe
  \item If NATed, a new relay reservation must be made
  \item All peer connections must be re-established from scratch
  \item Total recovery time: \textbf{10--30 seconds} vs iroh's \textbf{$<$5 seconds}
\end{enumerate}

\subsubsection{Moving Between WiFi Networks}

\begin{table}[H]
\centering
\small
\begin{tabular}{L{3.5cm} C{1.5cm} C{1.5cm} L{4cm}}
\toprule
\textbf{Scenario} & \textbf{LP2P} & \textbf{Iroh} & \textbf{LP2P Details} \\
\midrule
WiFi A $\to$ WiFi B (diff LAN) & \no$^*$ & \dgr & All connections lost. Full reconnection cycle. \\
\rowcolor{lightgray}
Same AP, IP unchanged & \yes & \yes & No disruption (identical). \\
WiFi A $\to$ brief gap $\to$ WiFi B & \no$^*$ & \dgr & LP2P: TCP/UDX timeout during gap. Iroh: QUIC may survive if within idle timeout. \\
\rowcolor{lightgray}
Same LAN, diff subnet & \no$^*$ & \dgr & LP2P: IP change = death. Iroh: QUIC migrates. \\
\bottomrule
\end{tabular}
\caption{WiFi roaming. $^*$LibP2P connections do not survive IP changes.}
\end{table}

\subsubsection{BLE Range Transitions}

BLE behavior is \textbf{identical} between libp2p and iroh---BLE is independent of the Internet transport. The only difference: when BLE drops and iroh/libp2p must take over, iroh's QUIC connection is more likely to still be alive (due to connection migration), while libp2p connections may have already been severed by a concurrent IP change.

\subsubsection{BLE-Only (No Internet)}

\textbf{Identical} to iroh. BLE-only operation does not use the Internet transport layer at all. All scenarios in Section 7.4 apply unchanged.

\subsection{Comparative Summary: LibP2P vs Iroh}

\begin{table}[H]
\centering
\small
\begin{tabular}{L{4cm} C{3cm} C{3cm} C{2.5cm}}
\toprule
\textbf{Dimension} & \textbf{LibP2P} & \textbf{Iroh} & \textbf{Winner} \\
\midrule
\multicolumn{4}{l}{\textbf{Architecture}} \\
\midrule
Identity model & Separate PeerId & NodeId = PubKey & Iroh \\
\rowcolor{lightgray}
Bootstrap required & Yes (critical) & No & Iroh \\
Address format & Multiaddr strings & IP:port + relay URL & Iroh \\
\rowcolor{lightgray}
Protocol & UDX (UDP) + TCP & QUIC (UDP) & Iroh \\
Local discovery & mDNS & None (BLE only) & LibP2P \\

\midrule
\multicolumn{4}{l}{\textbf{NAT Traversal}} \\
\midrule
Relay type & Peer-based (Circuit Relay v2) & Dedicated servers & Iroh \\
\rowcolor{lightgray}
Relay availability & Depends on peer topology & Managed infrastructure & Iroh \\
Hole punching & DCUtR (after relay) & Automatic (background) & Iroh \\
\rowcolor{lightgray}
NAT detection & AutoNAT v2 & Automatic & Tie \\
Cross-protocol (v4$\leftrightarrow$v6) & Requires dual-stack relay peer & Relay bridges natively & Iroh \\

\midrule
\multicolumn{4}{l}{\textbf{Mobility}} \\
\midrule
Connection migration & \textcolor{no}{\textbf{None}} & QUIC native & \textbf{Iroh} \\
\rowcolor{lightgray}
WiFi$\to$cellular recovery & 10--30s (full reconnect) & $<$5s (migration) & \textbf{Iroh} \\
WiFi roaming & Connection lost & QUIC survives & \textbf{Iroh} \\
\rowcolor{lightgray}
Train / frequent handoff & Unusable & Degraded but functional & \textbf{Iroh} \\

\midrule
\multicolumn{4}{l}{\textbf{Resilience}} \\
\midrule
Single point of failure & Bootstrap peer & Relay cluster & Iroh \\
\rowcolor{lightgray}
Self-hosted relay & Complex (run libp2p node) & Simple (iroh relay binary) & Iroh \\
Relay redundancy & Must configure multiple bootstrap & Multiple relay URLs & Iroh \\
\rowcolor{lightgray}
Offline operation (BLE) & Identical & Identical & Tie \\

\midrule
\multicolumn{4}{l}{\textbf{Where LibP2P Was Better}} \\
\midrule
Same-LAN discovery & mDNS (automatic) & None & \textbf{LibP2P} \\
\rowcolor{lightgray}
TCP fallback & Works through HTTP proxies & UDP only (QUIC) & \textbf{LibP2P} \\
Peer-as-relay & Any public peer can relay & Only dedicated servers & \textbf{LibP2P} \\
\rowcolor{lightgray}
Ecosystem interop & IPFS-compatible & Iroh-only & \textbf{LibP2P} \\
\bottomrule
\end{tabular}
\caption{Head-to-head comparison. Iroh wins on mobility, simplicity, and reliability. LibP2P had advantages in local discovery (mDNS) and TCP fallback.}
\end{table}

\subsection{Scenario-by-Scenario Comparison}

\begin{longtable}{L{3.5cm} L{2cm} C{1.5cm} C{1.5cm} L{3cm}}
\toprule
\textbf{Scenario} & \textbf{Addr Config} & \textbf{LP2P} & \textbf{Iroh} & \textbf{Key Difference} \\
\midrule
\endfirsthead
\toprule
\textbf{Scenario} & \textbf{Addr Config} & \textbf{LP2P} & \textbf{Iroh} & \textbf{Key Difference} \\
\midrule
\endhead
\midrule
\multicolumn{5}{r}{\textit{continued on next page}} \\
\bottomrule
\endfoot
\bottomrule
\endlastfoot

\multicolumn{5}{l}{\textbf{Co-located Peers}} \\
\midrule
Same WiFi, both NATed & PIv4 / PIv4 & \yes & \yes & LP2P: mDNS finds peer directly. Iroh: must use relay. \textbf{LP2P better.} \\
\rowcolor{lightgray}
Same WiFi, firewall blocks UDP & PIv4 / PIv4 & \yes & \no & LP2P: falls back to TCP. Iroh: QUIC is UDP-only. \textbf{LP2P better.} \\
Different WiFi, same cafe & PIv4 / PIv4 & \dgr & \dgr & Both need relay/NAT traversal. Similar. \\
\rowcolor{lightgray}
No Internet, both BLE & BLE / BLE & \yes & \yes & Identical. BLE only. \\

\midrule
\multicolumn{5}{l}{\textbf{Remote Peers}} \\
\midrule
Both home WiFi (NAT) & PIv4 / PIv4 & \dgr & \dgr & LP2P: bootstrap$\to$relay. Iroh: relay server. Similar quality. \\
\rowcolor{lightgray}
One public, one NATed & PuIv4 / PIv4 & \yes & \yes & Both handle this well. \\
Both public IPv6 & PuIv6 / PuIv6 & \yes & \yes & Both: direct. LP2P used UDX, iroh uses QUIC. \\
\rowcolor{lightgray}
IPv4 $\leftrightarrow$ IPv6 mismatch & PIv4 / PuIv6 & \dgr & \dgr & LP2P: needs dual-stack relay peer. Iroh: relay bridges natively. \textbf{Iroh slightly better.} \\
Bootstrap/relay down & PIv4 / PIv4 & \no & \dgr & LP2P: complete failure for NATed peers. Iroh: can try other relays. \textbf{Iroh better.} \\

\midrule
\multicolumn{5}{l}{\textbf{Mobility}} \\
\midrule
WiFi $\to$ cellular & PIv4 $\to$ PIv4 & \no & \dgr & LP2P: 10--30s blackout. Iroh: $<$5s. \textbf{Iroh far better.} \\
\rowcolor{lightgray}
Roaming between APs & PIv4 $\to$ PIv4 & \no & \dgr & Same. No connection migration in LP2P. \textbf{Iroh far better.} \\
On a train & PIv4 (changing) & \no & \dgr & LP2P: constant full reconnects. Iroh: QUIC adapts. \textbf{Iroh far better.} \\
\rowcolor{lightgray}
Walking away from BLE peer & PIv4 & \dgr & \yes & LP2P: if IP also changed, both transports lost. Iroh: QUIC survives. \\

\midrule
\multicolumn{5}{l}{\textbf{Edge Cases}} \\
\midrule
Corporate firewall (TCP only) & PIv4 & \yes & \no & LP2P: TCP transport works. Iroh: QUIC blocked. \textbf{LP2P better.} \\
\rowcolor{lightgray}
Restrictive mobile carrier (CGNAT + symmetric NAT) & PIv4 & \dgr & \dgr & Both relay-only. LP2P relay depends on bootstrap availability. Iroh relay is managed. \textbf{Iroh slightly better.} \\
Hostile network (UDP throttled) & PIv4 & \prt & \dgr & LP2P: TCP fallback. Iroh: degraded QUIC. \textbf{LP2P better.} \\

\end{longtable}

\subsection{What Was Lost in the Migration}

Three capabilities were lost when moving from libp2p to iroh:

\begin{enumerate}
  \item \textbf{mDNS local discovery}: LibP2P automatically discovered peers on the same LAN via mDNS broadcasts. With iroh, same-LAN peers must either be in BLE range or exchange addresses out-of-band. This is a real regression for the ``both on same WiFi but not in BLE range'' scenario.

  \item \textbf{TCP fallback}: LibP2P could fall back to TCP when UDP was blocked (corporate firewalls, some hotel networks). QUIC is UDP-only. If UDP is blocked, iroh cannot connect at all. This matters in restrictive network environments.

  \item \textbf{Peer-as-relay}: Any libp2p peer with a public IP could serve as a Circuit Relay. This provided organic relay infrastructure that grew with the network. With iroh, relay is centralized to iroh.computer's servers (or self-hosted). The network cannot organically grow its own relay capacity.
\end{enumerate}

\subsection{What Was Gained in the Migration}

\begin{enumerate}
  \item \textbf{Connection migration (QUIC)}: The single biggest improvement. Mobile users experience dramatically better connectivity during network transitions. This alone may justify the migration for a mobile-first application.

  \item \textbf{No bootstrap dependency}: Eliminating the bootstrap peer removed a critical single point of failure. The network no longer depends on one server being up.

  \item \textbf{Unified identity}: NodeId = public key eliminates the double-mapping complexity and reduces the attack surface for identity confusion.

  \item \textbf{Managed relay infrastructure}: Iroh's relay servers are operated professionally with geographic distribution, monitoring, and redundancy. The bootstrap peer was a single VM.

  \item \textbf{Simpler architecture}: Fewer moving parts (no AutoNAT, no AutoRelay, no DCUtR, no mDNS, no Circuit Relay). Fewer things that can break.

  \item \textbf{Better hole-punching}: Iroh's hole-punching runs continuously in the background and automatically migrates from relay to direct. LibP2P's DCUtR was a one-shot attempt after relay establishment.
\end{enumerate}

% =============================================================================
\section{Summary}
% =============================================================================

\begin{table}[H]
\centering
\begin{tabular}{L{5cm} C{2cm} C{2cm}}
\toprule
\textbf{Scenario Class} & \textbf{BLE} & \textbf{Iroh} \\
\midrule
Co-located, any network & \yes & Varies \\
Remote, both have routable IP & \na & \yes \\
Remote, one or both NATed & \na & \dgr \\
Remote, one or both BLE-only & \na & \no \\
Mobility (IP change) & Unaffected & \dgr \\
Mobility (BLE range change) & Varies & Unaffected \\
No Internet at all & \yes{} (if in range) & \no \\
\bottomrule
\end{tabular}
\caption{High-level connectivity summary.}
\end{table}

The two transports are \textbf{complementary}: BLE provides resilient local connectivity that is immune to network configuration, while iroh provides global reach that is immune to physical proximity. Together they cover most real-world scenarios. The primary gap---no relaying and no offline delivery---is a deliberate design choice that keeps the protocol simple, private, and predictable.

The migration from libp2p to iroh was a net positive for mobile-first use cases. The loss of mDNS and TCP fallback is real but narrow; the gain in connection migration (QUIC), architectural simplicity, and relay reliability is broad and impactful. The comparison table below distills the trade-off:

\begin{table}[H]
\centering
\begin{tabular}{L{5cm} C{2cm} C{2cm}}
\toprule
\textbf{Scenario Class} & \textbf{LibP2P} & \textbf{Iroh} \\
\midrule
Co-located, same LAN & \yes{} (mDNS) & \dgr{} (no mDNS) \\
Co-located, BLE range & \yes & \yes \\
Remote, both NATed & \dgr & \dgr \\
Remote, one public & \yes & \yes \\
Corporate firewall (TCP only) & \yes & \no \\
Mobility (IP change) & \no{} (reconnect) & \dgr{} (migrates) \\
Mobility (BLE transition) & Identical & Identical \\
Bootstrap/relay failure & \no{} (SPOF) & \dgr{} (redundant) \\
\bottomrule
\end{tabular}
\caption{LibP2P vs Iroh: high-level comparison by scenario class.}
\end{table}

\end{document}
